Simulation of 2D physics of hand drawn objects using OpenCV and Box2D

### 2011-02-07 latex/paper.tex
dalsia kapitola rozpisana

\documentclass{ifacconf}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{natbib}

@@

\section{Physics simulation in Box2D}
There is lot of physics engines that can be used for simulation of physics.
Because we wanted to simulate physics only in 2D we could, code our own
implementation of physics, or use one of commercial or open source engines. We
chose Box2D[\cite{GameEngines}], which is open source 2D physiscs engine with
posibility to simulate rigid body objects and their collisions.
\subsection{World}
To create physics simulation we need to create world. World is object that
manages memory, objects and simulation. Creation of world is shown in Listing
\ref{World}:
\begin{lstlisting}[caption=Creation of Box2D world,label=World]
self.worldAABB=box2d.b2AABB()
self.worldAABB.lowerBound = (-100.0, -100.0) %\label{World:lower})
self.worldAABB.upperBound = ( 600.0, 600.0)	%\label{World:upper})
gravity = (0.0, -10.0) %\label{World:gravity})

doSleep = True %\label{World:sleep})
self.world = box2d.b2World(self.worldAABB, gravity, doSleep)
\end{lstlisting}
First we have to create boundaries of the world. We define them as vectors from
bottom left (line: \ref{World:lower}) to top right (line:
\ref{World:upper}). Objects have to be inside the boudaries, when an object
touch the boundary it gets stuck. Then we define gravity vector (line:
\ref{World:gravity}). The last thing before creation of the world we allow
objects to sleep(line: \ref{World:sleep}). Object that are not moving fall
asleep, then are ignored by the engine. Last line of Listing \ref{World} creates
the world.

When we have world created we are ready to create objects from outlines.
\subsection{Objects}
Every object that is simulated in Box2D consists of body and shapes.
\subsubsection{Bodies, shapes and collisions}
Bodies are skelet used by shapes. One body can contain more shapes, but one
shape could be attached to only one body. Box2d is rigid body
physics engine, that mean that shapes attached to body can not move against
other, or body.
\subsection{Tesselation}
Box2D supports only collisions between convex objects. That is why we need to
breake outlines to convex polygons. There is more ways how to break concave
objects. We chose Seidel's Triangulation Algorithm implemented by poly2tri
Python library
%TODO asi iny algorytmus
%TODO pridat referenciu na poly2tri

@@

\section*{Acknowledgments}
The work was supported by a grant (No. NIL-I-007-d) from Iceland, Liechtenstein 
and Norway through the EEA Financial Mechanism and the Norwegian Financial 
Mechanism. This project is also co-financed from the state budget of the Slovak
Republic..

>>>>

\documentclass{ifacconf}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{natbib}

@@

\section{Physics simulation in Box2D}
There is lot of physics engines that can be used for simulation of physics.
Because we wanted to simulate physics only in 2D we could, code our own
implementation of physics, or use one of commercial or open source engines. We
chose Box2D[\cite{GameEngines}], which is open source 2D physiscs engine with
possibility to simulate rigid body objects and their collisions.
\subsection{World}
To create physics simulation we need to create world. World is object that
manages memory, objects and simulation. Creation of world is shown in Listing
\ref{World}:
\begin{lstlisting}[caption=Creation of Box2D world,label=World]
self.worldAABB=box2d.b2AABB()
self.worldAABB.lowerBound = (-100.0, -100.0) %\label{World:lower})
self.worldAABB.upperBound = ( 600.0, 600.0)	%\label{World:upper})
gravity = (0.0, -10.0) %\label{World:gravity})

doSleep = True %\label{World:sleep})
self.world = box2d.b2World(self.worldAABB, gravity, doSleep)
\end{lstlisting}
First we have to create boundaries of the world. We define them as vectors from
bottom left (line: \ref{World:lower}) to top right (line:
\ref{World:upper}). Objects have to be inside the boundaries, when an object
touch the boundary it gets stuck. Then we define gravity vector (line:
\ref{World:gravity}). The last thing before creation of the world we allow
objects to sleep(line: \ref{World:sleep}). Object that are not moving fall
asleep, then are ignored by the engine. Last line of Listing \ref{World} creates
the world.

World is created and we are ready to create objects from outlines.
\subsection{Objects}
Every object that is simulated in Box2D consists of body and shapes. Our objects
are described by the contour tree. To create objects we need to iterate contour
tree, find contours that belongs to together and create objects for these
contours.
\subsubsection{Contour tree}
Contour tree is object in which are stored points of each contour. Contours are
connected by functions returning reference to other contours with h\_next() and
v\_next(). h\_next() is referencing to deeper contour, and v\_next is
referencing to another object contour. To iterate over all contours we have
created recursive function shown in Listing: \ref{CreateObjects}.
\begin{lstlisting}[label=CreateObjects,caption=Function to iterate through
contour tree]
def CreateObjectsFromCountours(self,cont,h=0,v=0):
  if v>0:	%\label{CreateObjects:depth})
    density = 10.0 %\label{CreateObjects:density10})
  else:
    density = 0 %\label{CreateObjects:density0})
  if len(cont)>8:
    self.CreateObject(cont,h,v)

  if cont.v_next():
    v += 1
    self.CreateObjectsFromCountours(cont.v_next(),h,v)
    v -= 1

  if cont.h_next():
    h += 1
    self.CreateObjectsFromCountours(cont.h_next(),h,v)
\end{lstlisting}
We iterate through contour tree. First level is outer contour of the image
(Line: \ref{CreateObjects:depth}), because we do not want the outer contour to
move, we set it as static byt setting density to 0 (Line:
\ref{CreateObjects:density0}). Every other contour is dynamic body with density
set to 10 (Line: \ref{CreateObjects:density10}).

When we know what type of object we will create, we can create bodies and
shapes for our contours.
\subsubsection{Bodies, shapes and collisions}
Bodies are backbone used by shapes. One body can contain more shapes, but one
shape could be attached to only one body. Box2d is rigid body physics engine,
that mean that shapes attached to body can not move against other, or body. Body
have position and velocity. Forces, torques and impulses can be applied to body
[\cite{Box2DManual}].  Bodies just hold the shapes. Shapes are elements that
collide together.

Listing \ref{Objects} shows how to create object from outer contour: 
\begin{lstlisting}[name=Objects,label=Objects,caption=Creation of object from
outer contour]
def CreateObject(self, cont, h,v):
  contM = []
  for point in cont:
    x = point[0]/30.0	%\label{Objects:scaling})
    y = point[1]/30.0
    contM.append((x,y))

  bd=box2d.b2BodyDef() %\label{Objects:CreateBody})
  bd.position = ( 0.0, 0.0 )

  edgeDef=box2d.b2EdgeChainDef() %\label{Objects:CreateChain})
  edgeDef.setVertices(contM) %\label{Objects:AssignVertices})

  if v==0:
    body = self.world.CreateBody(bd)
    try:
      self.contourBodies.append(body)
    except:
      self.contourBodies = [body] %\label{Objects:CountourBodies})
      body.CreateShape(edgeDef) %\label{Objects:AttachShape})
\end{lstlisting}
Image size is measured in pixels and Box2D units are kilograms, meters,
and seconds (KMS) we should scale images coordinates to fit in 0.1m to
10m. In that scale is performance of Box2D the best. We are doing it by
dividing of value of pixel coordinates by 30.0 (Line: \ref{Objects:scaling})

Then we create a body definition that will represent our contour(Line:
\ref{Objects:CreateBody}) and set up it initial possition in next line.

After that we create shape of body as chain of edges (Line:
\ref{Objects:CreateChain}) and assign the array of vertices to it (Line:
\ref{Objects:AssignVertices}). Edges are special type of shapes that have no
mass, is represented as line betven vertices that collide with other non-edge
objects. Edges are easy to create because they do not have to be concave unlike
polygons.

At last we attach this shape to created body \ref{Objects:AttachShape}. Because
Box2D does not keep track about body definitions, we have to store bodies in to
array for later use\ref{Objects:CountourBodies}.

Listing of the function CreateObject() continous in Listing \ref{Objects}. This
part of function creates dynamic objects inside the outer contour.
\begin{lstlisting}[name=Objects,firstnumber=auto,label=Objects,caption=Creation of objecs] if v == 1: try:
      body = self.objectBodies[h]
    except:
      body = self.world.CreateBody(bd)
      self.objectBodies[h] = body
    polyline = []
    for (x,y) in cont:
      polyline.append(p2t.Point(x,y))
    cdt = p2t.CDT(polyline)
    triangles = cdt.triangulate()
    for t in triangles:
      x1 = t.a.x/30.0
      y1 = t.a.y/30.0
      x2 = t.b.x/30.0
      y2 = t.b.y/30.0
      x3 = t.c.x/30.0
      y3 = t.c.y/30.0
      if math.hypot(x2-x1,y2-y1)<0.1:
        x2 = x2 + math.copysign(0.1, x2-x1)
        y2 = y2 + math.copysign(0.1, y2-y1)
        print "kratky vektor"
      if math.hypot(x3-x2,y3-y2)<0.1:
        x3 = x3 + math.copysign(0.1, x3-x2)
        y3 = y3 + math.copysign(0.1, y3-y2)
        print "kratky vektor"
      if math.hypot(x1-x3,y1-y3)<0.1:
        x1 = x1 + math.copysign(0.1, x1-x3)
        y1 = y1 + math.copysign(0.1, y1-y3)
        print "kratky vektor"
      poly=box2d.b2PolygonDef()
      poly.setVertices(((x1, y1), (x2, y2), (x3, y3)))
      poly.density = 1.0
      poly.restitution = 0.0
      poly.friction = 0.0
      body.CreateShape(poly)
      body.SetMassFromShapes()
\end{lstlisting}

\subsection{Tesselation}
Box2D supports only collisions between convex objects. That is why we need to
break outlines to convex polygons. There is more ways how to break concave
objects. We chose the 2D constrained Delaunay triangulation algorithm
implemented by poly2tri Python library[\cite{Delauanay}].

@@

\section*{Acknowledgments}
The work was supported by a grant (No. NIL-I-007-d) from Iceland, Liechtenstein 
and Norway through the EEA Financial Mechanism and the Norwegian Financial 
Mechanism. This project is also co-financed from the state budget of the Slovak
Republic.
